An asymptotically normal out-of-sample test of equal predictive accuracy for nested models

### 2012-03-30 Paper.tex
Drafted initial version of the paper

\usepackage{amsfonts}
\usepackage{amsmath,amsthm,amssymb,graphicx,setspace,url,booktabs,tabularx}
\usepackage[sort,round]{natbib}
\usepackage[margin=1.25in]{geometry}
\usepackage[small]{caption}

\bibliographystyle{abbrvnat}
\newcommand\citepos[2][]{\citeauthor{#2}'s \citeyearpar[#1]{#2}}
\newcommand\poscw{\citeauthor{ClW:06}'s \citeyearpar{ClW:06,ClW:07}}
\newcommand\citen[1]{\citeauthor{#1}, \citeyear{#1}}
\frenchspacing
\onehalfspacing

\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{claim}[thm]{Claim}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lema}{Lemma}[section]
\newtheorem{alg}{Algorithm}
\newtheorem{asmp}{Assumption}[section]

\theoremstyle{definition}

\newtheorem{example}{Example}
\newtheorem{defn}{Definition}
\newtheorem{rem}{Remark}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}
%\DeclareMathOperator{\vec}{vec}
\DeclareMathOperator{\vech}{vech}

\DeclareMathOperator{\pr}{Pr}

\newcommand{\X}{\ensuremath{\mathrm{X}}}
\newcommand{\R}{\ensuremath{\mathrm{R}}}
\newcommand{\p}{\ensuremath{\mathrm{P}}}

\newcommand{\dd}[1]{\frac{\partial}{\partial #1}}
\newcommand{\ned}{\textsc{ned}}
\newcommand{\clt}{\textsc{clt}}
\newcommand{\fclt}{\textsc{fclt}}
\newcommand{\aic}{\textsc{aic}}
\newcommand{\bic}{\textsc{bic}}
\newcommand{\dgp}{\textsc{dgp}}
\newcommand{\ols}{\textsc{ols}}
\newcommand{\mds}{\textsc{mds}}
\newcommand{\oos}{\textsc{oos}}
\newcommand{\gdp}{\textsc{gdp}}
\newcommand{\hac}{\textsc{hac}}
\newcommand{\fwe}{\textsc{fwe}}
\newcommand{\cdf}{\textsc{cdf}}
\newcommand{\ma}{\textsc{ma}}

\renewcommand{\Re}{\ensuremath{\mathbb{R}}}

@@

\author{Gray Calhoun\thanks{Address: 467 Heady Hall, Economics
    Department; Iowa State University; Ames, IA 50011. Telephone:
    (515) 294-6271.  Email: \texttt{gcalhoun@iastate.edu}.} \\ Iowa
  State University}

\title{An asymptotically normal out-of-sample
  test of equal predictive accuracy for nested models} \date{\today}
\maketitle

\begin{abstract} 
  \noindent This paper proposes a modification of \citepos[J.
  Econom.]{ClW:07} adjusted out-of-sample $t$-test.  The
  alternative model is still estimated with a fixed-length rolling
  window, but the benchmark is estimated with a recursive window. The
  resulting statistic is asymptotically normal even when the models
  are nested.  Moreover, the alternative model can be estimated using
  popular model selection methods, such as the \aic\ or \bic.  This
  paper also presents a new bootstrap procedure based on
  \citet[Econometrica]{RoW:05} to compare multiple models while
  controlling family-wise error, and demonstrates this procedure by
  analyzing \citepos[Rev. Finan. Stud.]{GoW:08} excess returns dataset.

@@

\noindent Keywords: Forecast Evaluation, Martingale Difference
Sequence, Model Selection, Family-wise Error; Multiple Testing;
Bootstrap; Reality Check

@@

\newpage \noindent In a pair of papers, \citet{ClW:06,ClW:07} develop
an out-of-sample (\oos) test of the null hypothesis that a small
benchmark model is correctly specified.  Their test compares the
forecasting performance of a pair of nested models, and the null
hypothesis is that the innovations in the smaller model form a
martingale difference sequence.  This test procedure is
popular,\footnote{As of 03 June 2011, \citet{ClW:06} has been cited by
  26 papers indexed by the \textsc{isi} Web of Knowledge, and
  \citet{ClW:07} by 37. They have been cited by 113 and 165 papers
  respectively indexed by Google Scholar.} and one assumes that this
is due in part to the statistic's convenience.  The statistic is
approximately normal after adjusting for the estimation error of the
larger model.  Normality comes from a fixed-length rolling window, as
in \citet{GiW:06}, and the adjustment centers the statistic at the
appropriate population quantities.  This statistic is especially
convenient because other tests for similar hypotheses \citep[among
others]{Mcc:07,ClM:05,ClM:01,CoS:04,CoS:02,CCS:01} have a nonstandard
limit distribution and place restrictions on the models under
consideration, while other asymptotically normal statistics test a
different null hypothesis \citep{GiW:06} or place assumptions on the
models and \dgp\ that are rarely appropriate in empirical work
\citep{Wes:96,WeM:98,Mcc:00,DiM:95}.

However, Clark and West's statistic is only ``approximately normal''
in an informal sense.  Clark and West present Monte Carlo evidence of
the statistic's distribution, but only prove that the statistic is
asymptotically normal when the benchmark model is a random walk
\citep{ClW:06}. Estimating the parameters of the smaller model
invalidates their proof.

In this paper, I prove that a modified version of their statistic is
asymptotically normal even when the smaller model is estimated.  To
achieve normality, we need a consistent estimate of the pseudo-true
benchmark model, while maintaining an inconsistent estimate of the
larger model so that we can ignore nesting.  We can meet both needs by
using different window strategies for each model; the benchmark model
is estimated using a recursive window and the alternative with a
fixed-length rolling window.

Mixing window strategies is uncommon but needn't be. In most
applications, the null hypothesis imposes both equal accuracy between
the two models and stability.  The benchmark model rarely allows for
breaks, parameter drift, or other forms of
instability,\footnote{Exceptions are \citepos{StW:07}
  \textsc{ima}(1,1) and \textsc{uc-sv} models of inflation.} but the
researcher is typically concerned about instability.  Indeed, concern
about instability is often given as a reason for doing an \oos\
analysis, especially with a short rolling window.\footnote{This
  motivation is discussed in \citet{StW:03}, \citet{PeT:05,PeT:07},
  \cite{GiW:06}, \citet{GoW:08}, \citet{ClM:09c}, and
  \cite{GiR:09,GiR:10}, among others.} A researcher could impose
stability on both models by using a recursive window or relax
stability for both by using a rolling window; both approaches should
not affect the test's size, but may affect power.  But the researcher
could instead impose stability on the benchmark and relax it for the
alternative by using a recursive window for the benchmark and a
rolling window for the alternative model.  This approach could have a
power advantage and is similar in spirit to using a likelihood ratio
test instead of an \textsc{lm} or Wald test.

This statistic has a substantial advantage over existing \oos\
statistics for nested models: the alternative can be essentially
arbitrary, as long as it obeys the necessary moment conditions.  In
particular, researchers can use model selection techniques like the
\aic\ or \bic\ to determine the number of lags to include, the
particular exogenous variables to include, etc.  Other methods that
test a similar hypothesis are unable to handle these models
\citep[except][which does not allow the benchmark to be
estimated]{ClW:06}; \citet{GiW:06} are able to handle such models for
both the alternative and the benchmark but, as mentioned earlier, they
test a different aspect of forecasting performance.

This paper focuses on nested models, as they have received the most
attention in the empirical and theoretical literature, but the
statistic can be used with non-nested models as well.  This generality
is useful, since \citepos{Wes:96} results do not apply to non-nested
models if they both encompass the true \dgp, which is allowable under
the null: in the limit, both models will converge to the \dgp\ and
give identical forecasts.  Consequently, the naive \oos\ $t$-test is
invalid, even after correcting the standard error to reflect parameter
uncertainty.  Constructing an \oos\ test for non-nested models using
\citepos{ClM:01} approach would be difficult.  To my knowledge, this
paper gives the first \oos\ test of equal predictive ability for
non-nested models that does not implicitly rule out both models
encompassing the \dgp.\footnote{Again, \citet{GiW:06} does not make
  this implicit assumption, but their statistic tests a different null
  hypothesis.}

Statistics for comparing two models are useful, but a researcher often
has a set of potential models, and wants to know which of them
significantly outperform the benchmark.  As \citet{Whi:00} shows,
looking at the naive $p$-values is misleading, but bootstrapping the
statistics can give a valid test (White calls this procedure the
``bootstrap reality check'').  So I present a test for equal
predictive ability of multiple models based on \citepos{RoW:05} StepM,
which uses a step-down procedure that iteratively rejects models to
achieve higher power than the reality check (see Theorem~\ref{res:2}
for details) and will indicate which of the models improves on the
benchmark.

I also present a new result for the validity of using the stationary
boostrap with \oos\ statistics, which is necessary to verify that the
StepM is appropriate.  Existing results on boostrapping \oos\
statistics have some drawbacks.  \citet{Whi:00} and \citet{Han:05} use
the stationary bootstrap, but require the test sample to be much
smaller than the training sample, which is obviously inappropriate
here.  \citet{CoS:07} relax that requirement, but only consider the
moving block bootstrap \citep{LiS:92,Kun:89} and make the statistic
more complicated than necessary by adjusting the objective function of
the bootstrapped statistic to center it correctly.\footnote{Their
  recentering is required for consistency and does not give higher
  order accuracy.}  A parametric bootstrap, such as that used by
\citet{ClM:11}, would be difficult to implement here since we want to
test whether several different null hypotheses hold individually (as
I'll discuss later, Clark and McCracken test a single null hypothesis
that describes the behavior of all of the models under consideration).
So this paper's result for the stationary bootstrap of \oos\ statistics
improves on existing procedures.

The next section presents our theoretical result and
Section~\ref{sec:2} presents two simulations that compare our pairwise
\oos\ test to \poscw\ original statistic.  Section~\ref{sec:3}
demonstrates the use of our statistic by reanalyzing \citepos{GoW:08}
study of excess return predictability. Section~\ref{sec:4} concludes.

\section{Theoretical Results}\label{sec:1}
The first result modifies \citepos{ClW:07} so that it is
asymptotically normal. The conditions are essentially the same in
existing papers
\citep[e.g.][]{ClW:07,ClW:06,Wes:96,WeM:98,Mcc:00,GiW:06}, but with a
modified window scheme.  Our result follows from \citet{Wes:96},
treating the alternative model's forecasts as additional data. Note
that asymptotic equivalence does not hold so we need to account for
parameter estimation error.

\begin{thm}\label{res:1}
  Suppose that we have two models $\hat{y}_{0t}$ and $\hat{y}_{1t}$ to
  forecast the variable $y_t$, and have observations for
  $t=1,\dots,T+1$.  Assume the following hold:
  \begin{enumerate}
  \item \label{item:1} The benchmark forecast $\hat{y}_{0,t+1}$, is
    estimated using \ols\ with a recursive window: $\hat{y}_{0,t+1} =
    x_t'\hat{\beta}_t$ for some vector of predictors $x_t$ with
    \begin{equation}
      \hat{\beta}_t = \Big(\sum_{s=2}^{t} x_{t-1} x_{t-1}'\Big)^{-1}
      \sum_{s=2}^t x_{t-1} y_t, \qquad t = R,\dots,T.
    \end{equation}
    Also define $\beta_0 = \big(\sum_{t=2}^{T} \E x_{t-1}
    x_{t-1}'\big)^{-1} \sum_{t=2}^T \E x_{t-1} y_t$.
  \item \label{item:2} The alternative forecast, $\hat{y}_{1t}$, is
    estimated using a rolling window of fixed length $R$ (which is
    less than $T$), so $\hat{y}_{1,t+1} =
    \psi(y_t,z_t,\dots,y_{t-R+1}, z_{t-R+1})$ where $\psi$ is a known
    function and $z_t$ is a sequence of predictors that includes
    $x_t$.
  \item \label{item:3} The series $y_t$ and $z_t$ are \ned\ of size
    $-1/2$ on a strong mixing series of size $-r/(r-2)$ for $r>2$.
    Also, $y_t$, $\hat{y}_{1t}$, and $x_t$ have uniformly bounded $2
    r+\delta$ moments for some positive $\delta$.
  \item \label{item:4} Define \[f_t(\beta) = (y_{t+1} - x_t'\beta)^2 -
    (y_{t+1} - \hat{y}_{1,t+1})^2 + (x_t'\beta - \hat{y}_{1,t+1})^2,\]
    $\bar{f}(\beta) = P^{-1} \sum_{t=R+1}^T f_t(\beta)$, and $\bar f =
    P^{-1} \sum_{t=R+1}^{T} f_t(\hat{\beta}_t)$, with $P = T - R$. The
    processes $f_t(\beta_0)$ and $x_t(y_{t+1} - x_t'\beta)_0$ are
    covariance stationary and $\bar f(\beta_0)$ has uniformly positive
    long run variance.
  \end{enumerate}
  Under the null hypothesis that $y_t - \hat{y}_{0t}(\beta_0)$ is a
  martingale difference sequence with respect to the filtration
  $\mathcal{F}_t = \sigma(y_t, z_{t-1}, y_{t-1}, z_{t-2},\dots)$, $\sqrt{P} \bar f /
  \hat{\sigma} \to^d N(0,1)$, where
  \begin{align*}
    \hat{\sigma}^2 &= \hat{S}_{ff} + 2 \Pi (\hat{S}_{fg} + \hat{S}_{gg}), &
    \hat{S}_{ff} &= \frac1P \sum_{t=R+1}^T (f_t(\hat{\beta}_t) - \bar
    f)^2, \\
    \hat{S}_{fg} &= \frac1P \sum_{t=R+1}^T (f_t(\hat{\beta}_t) -
    \bar{f})(g_t(\hat{\beta}_t) - \bar g)', &
    \hat{S}_{gg} &= \frac1P \sum_{t=R+1}^T (g_t(\hat{\beta}_t) - \bar
    g)(g_t(\hat{\beta}_t) - \bar g)',
  \end{align*}
  $\Pi = 1 - (R/P) \log(1 + P/R)$, $\bar{g} = T^{-1}
  \sum_{t=R+1}^T g_t(\hat{\beta}_t)$, $\X' = [x_1,\dots,x_T]$, and
  \begin{equation*}
  g_t(\beta) = \Bigg\{\frac{2}{P}\sum_{t=R}^{T-1} x_t (2 x_t'\beta -
    y_{t+1} - \hat{y}_{1,t+1}) \Bigg\} \big(\X'\X / T \big)^{-1}  x_t(y_{t+1} - x_t'\beta).
  \end{equation*}
\end{thm}

The following remarks are relevant to Theorem~\ref{res:1}:

\begin{rem}
  Forecasters will almost always be interested in the one-sided
  alternative that $\E f_t(\beta_0) > 0$; i.e. that the alternative
  model is expected to forecast better than the benchmark.
\end{rem}

@@

\begin{rem}
  The requirement that the asymptotic variance of $\bar f(\beta_0)$ be
  uniformly positive is much less restrictive than in \cite{Wes:96}.
  As in \cite{GiW:06} and \citet{ClW:06,ClW:07}, the assumption only
  serves to rule out pathological cases---for example, letting both
  the benchmark and the alternative model be white noise. In
  \citet{Wes:96}, this assumption is a restriction on the \dgp\ as
  well as the forecasting models, but in this paper it is a
  restriction only on the models.
\end{rem}

\begin{rem}
  As in \citet{Wes:96}, $\Pi \to 1$ as $T \to \infty$ since $R$ is
  fixed.  But using West's general formula for $\Pi$ (which holds when
  $P/R$ converges as $T \to \infty$) can improve the statistic in
  practice, since researchers often invoke ``fixed $R$'' asymptotics
  when $R$ is large.  Section~\ref{sec:2} shows via Monte Carlo that
  this approximation can be accurate even when $R$ and $P$ are equal.
\end{rem}

\begin{rem}
  If one wants to test the less restrictive null hypothesis that
  $y_{t1} - \hat{y}_{0t}$ is uncorrelated with $\hat{y}_{1t}$, one can
  replace $\hat{S}_{ff}$, $\hat{S}_{fg}$ and $\hat{S}_{gg}$ with \hac\
  counterparts.  Lemma~\ref{res:a5} presents a more general version of
  the theorem that can cover this case.
\end{rem}

\begin{rem}
  The statistic we present tests the null hypothesis that the forecast
  errors from the population version of the benchmark model are a
  martingale difference sequence.  This may not be appropriate,
  depending on the loss function or utility function of interest.  Our
  statistic can be modified to test implications of optimal forecasts
  under other loss functions \citep[see][]{PaT:07,PaT:07b}.  Again,
  Lemma~\ref{res:a5} can cover these other applications.
\end{rem}

Before presenting the method for testing multiple models, I'll give a
version of \citepos{PoR:94} stationary bootstrap for the \oos\
statistic of Theorem~\ref{res:1}.  Note that the result holds as well
for the moving blocks bootstrap of \citet{Kun:89} and \citet{LiS:92}.
This result will be used in Theorem~\ref{res:2} to derive the joint
distribution of several models' \oos\ statistics.  The
boostrap-induced probability measure is denoted $\pr^{*}$.

\begin{lem}\label{res:3}
  Suppose the conditions of Theorem~\ref{res:1} hold, but strengthen
  condition~\ref{item:3} so that $y_t$ and $z_t$ are \ned\ of size
  $-1$ on a strong mixing series of size $-(2r + \rho)/(r-2)$ with
  $\rho > 0$ and $r > 2$.  Consider the following procedure:
  \begin{enumerate}
  \item Define the data matrices $\X_\R = (X_1,\dots,X_R)'$ and $\X_{\p} =
    (X_{R+1},\dots,X_{T-1})$ with
    \begin{equation*}
      X_t = \begin{cases}
        (y_{t+1}, x_t')' & t \leq R \\
        (y_{t+1}, x_t', \hat{y}_{1,t+1})' & t > R.
      \end{cases}
    \end{equation*}
  \item Draw $B$ samples of $P$ observations from $\X_{\p}$ using the
    stationary bootstrap with expected block size $b$ and denote each
    sample as $\X_{\p, l}^{*}$; let $\X_l^{*} = [\X_{\R}', \X_{\p,l}^{*\prime}]'$.
  \item Estimate $\bar{f}^{*}_l$ and $\hat{\sigma}_l^{*}$ as in
    Theorem~\ref{res:1} for each bootstrap sample $\X^{*}_l$.
  \end{enumerate}

  Let $(\bar{f}^{*},\hat{\sigma}^{*})$ be random variables with the
  same (conditional) distribution as $(\bar{f}_l^{*},
  \hat{\sigma}^{*}_l)$.  If $b \to \infty$ and $b^2/P \to 0$
  \begin{gather}
    \pr\big[\sup_x \big| \pr^{*}\big[\sqrt{P}(\bar{f}^{*} -
    \bar{f}(\hat{\beta}_{T+1})) \leq x \big] - \pr\big[\sqrt{P}
    (\bar{f} - \E \bar{f}(\beta_0)) \leq x\big] \big| > \epsilon\big]
    \to 0 \intertext{and} \pr\Bigg[\sup_x \Bigg|
    \pr^{*}\Bigg[\frac{\bar{f}^{*} -
      \bar{f}(\hat{\beta}_{T+1})}{\hat{\sigma}^{*}/\sqrt{P}} \leq x
    \Bigg] - \pr\Bigg[\frac{\bar{f} - \E
      \bar{f}(\beta_0)}{\hat{\sigma}/\sqrt{P}} \leq x\Bigg] \Bigg| >
    \epsilon\Bigg] \to 0
  \end{gather}
  for all $\epsilon>0$.
\end{lem}

\begin{rem}
  \citet{Whi:00}, \citet{Han:05}, and \citet{CoS:07} use the
  distribution of $\sqrt{P}(\bar{f}^{*} - \bar{f})$ to approximate
  that of $\sqrt{P}(\bar{f} - \E \bar{f}(\beta_0))$, while
  Lemma~\ref{res:3} uses the distribution of $\sqrt{P}(\bar{f}^{*} -
  \bar{f}(\hat{\beta}_{T+1}))$; $\hat{\beta}_{T+1}$ plays the same
  role as $\beta_0$ under the bootstrap-induced probability measure.
  Since $\hat{\beta}_t - \hat{\beta}_{T+1} = O_p(\sqrt{T-t})$, using
  their bootstrap distribution is inconsistent unless $P/T$ vanishes
  asymptotically, as in White and Hansen's
  papers\footnote{\citet{Whi:00} and \citet{Han:05} actually require
    that $P$ grow a little bit slower than $o(T)$.} or unless
  $\hat{\beta}_t^{*}$ is further modified, as in Corradi and Swanson's
  paper.
\end{rem}

\begin{rem}
  It is likely that Lemma~\ref{res:3} holds even if $\sqrt{P} \bar{f}$
  is not asymptotically normal, but I leave that case to future
  research as it's not necessary for this paper's results.
\end{rem}

\begin{rem}
  How we handle the first $R$ observations does not affect the
  asymptotic properties since $R$ is finite.  I've chosen to put them
  at the beginning of each bootstrap sample so that sample means
  correspond to population means of the bootstrap distribution.  Note
  that the alternative forecasts are not defined for the first $R$
  observations, so running the stationary bootstrap over the full
  dataset is not possible.
\end{rem}

\begin{rem}
  See \citet{RoW:05,RoW:06} for guidance on block length
  selection.
\end{rem}

Finally, I use Lemma~\ref{res:3} to test multiple models at once.
There are two approaches one can take.  The first is to test whether
any of the models can improve on the benchmark; \citet{Whi:00},
\citet{Han:05}, and \citet{ClM:11} take this approach.  This approach
requires testing a single null hypothesis, that the best model is no
better than the benchmark, and one wants a test statistic to control
size as usual.  The second approach, taken by \citet{RoW:05}, is to
try to determine which models outperform the benchmark, and to find as
many as possible that do.  Now there are as many hypotheses as there
are models, so we need to control a quantity similar to size, the
family-wise error rate, which is the probability that one or more true
hypotheses is rejected.  To appreciate the difference between these
two approaches, suppose that there are two models and the first
outperforms a benchmark but the second doesn't.  Under the first
approach, finding that both models outperform the benchmark is fine:
the null (that none of them do) is false so we want to reject.  But
under the second approach, this finding is a failure, as we
incorrectly believe that the second model beats the benchmark.

While methods like \citepos{Whi:00} and \citepos{Han:05} control the
family-wise error, \citepos{RoW:05} StepM procedure will typically find
more outperforming models while still controlling the family-wise error
rate.  Theorem~\ref{res:2} shows how the \oos\ bootstrap of
Lemma~\ref{res:3} can be used with the StepM procedure.

>>>>

\usepackage{amsfonts}
\usepackage{amsmath,amsthm,amssymb,graphicx,setspace,url,booktabs,tabularx}
\usepackage[sort,round]{natbib}
\usepackage[margin=1in]{geometry}
\usepackage[small]{caption}

\bibliographystyle{abbrvnat}
\newcommand\possessivecite[1]{\citeauthor{#1}'s \citeyearpar{#1}}
\newcommand\poscw{\citeauthor{ClW:06}'s \citeyearpar{ClW:06,ClW:07}}
\newcommand\citen[1]{\citeauthor{#1} \citeyear{#1}}
\frenchspacing
\onehalfspacing

\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{claim}[thm]{Claim}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{asmp}{Assumption}
\newtheorem{example}{Example}
\newtheorem{defn}{Definition}
\newtheorem{rem}{Remark}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{E}
\newcommand{\dd}[1]{\frac{\partial}{\partial #1}}
\newcommand{\aic}{\textsc{aic}}
\newcommand{\bic}{\textsc{bic}}
\newcommand{\dgp}{\textsc{dgp}}
\newcommand{\ols}{\textsc{ols}}
\newcommand{\mds}{\textsc{mds}}
\newcommand{\oos}{\textsc{oos}}
\newcommand{\gdp}{\textsc{gdp}}
\newcommand{\hac}{\textsc{hac}}
\newcommand{\ma}{\textsc{ma}}

@@

\author{Gray Calhoun} \title{An asymptotically normal out-of-sample
  test\\ of equal predictive accuracy for nested models} 
\date{\today}
\maketitle

\begin{abstract} 
  \noindent This paper proposes a modification of
  \possessivecite{ClW:07} adjusted out-of-sample $t$-test. We continue
  to estimate the alternative model under a fixed-length rolling
  window, but estimate the benchmark model under a recursive
  window. The resulting statistic gives an asymptotically normal test
  statistic that the benchmark model is correctly specified, even when
  it is nested in the alternative.  Moreover, the alternative model
  can be estimated using popular model selection methods, such as the
  \aic\ or \bic.

@@

\noindent Keywords: Out-of-Sample, Forecast Evaluation, Martingale
Difference Sequence, Model Selection

@@

\newpage \noindent In a pair of papers, \citet{ClW:06,ClW:07} develop
an out-of-sample (\oos) test of the null hypothesis that a small
benchmark model is correctly specified.  Their test compares the
forecasting performance of a pair of nested models, and the null
hypothesis is that the innovations in the smaller model form a
martingale difference sequence.  This test procedure is
popular,\footnote{As of 03 June 2011, \citet{ClW:06} has been cited by
  26 papers indexed by the \textsc{isi} Web of Knowledge, and
  \citet{ClW:07} by 37. They have been cited by 113 and 165 papers
  respectively indexed by Google Scholar.} and one assumes that this
is due in part to the statistic's convenience.  The statistic is
approximately normal after adjusting for the estimation error of the
larger model.  Normality comes from a fixed-length rolling window, as
in \citet{GiW:06}, and the adjustment centers the statistic at the
appropriate population quantities.  This statistic is especially
convenient because other tests for similar hypotheses \citep[among
others]{Mcc:07,ClM:05,ClM:01,CoS:04,CoS:02,CCS:01} have a nonstandard
limit distribution and place restrictions on the models under
consideration, while other asymptotically normal statistics
\citep{GiW:06} test a different null hypothesis.

However, Clark and West's statistic is only ``approximately normal''
in an informal sense.  Clark and West present Monte Carlo evidence of
the statistic's distribution, but only prove that the statistic is
asymptotically normal when the benchmark model is a random walk
\citep{ClW:06}. Estimating the parameters of the smaller model
invalidates the proof.

In this paper, I prove that a slightly modified version of their
statistic is asymptotically normal even when the smaller model is
estimated.  To achieve normality, we need a consistent estimate of the
pseudo-true benchmark model, while maintaining an inconsistent
estimate of the larger model so that we can ignore nesting.  We can
meet both needs by using different window strategies for each model;
the benchmark model is estimated using a recursive window and the
alternative with a fixed-length rolling window.

Mixing window strategies is uncommon but needn't be. In most
applications, the null hypothesis imposes both equal accuracy between
the two models and stability.  The benchmark model rarely allows for
breaks, parameter drift, or other forms of
instability,\footnote{Exceptions are \possessivecite{StW:07}
  \textsc{ima}(1,1) and \textsc{uc-sv} models of inflation.} but the
researcher is typically concerned about instability.  Indeed, concern
about instability is often given as a reason for doing an \oos\
analysis, especially with a short rolling window.\footnote{This is
  discussed in \citet{StW:03}, \citet{PeT:05,PeT:07}, \cite{GiW:06},
  \citet{GoW:08}, \citet{ClM:09c}, and \cite{GiR:09,GiR:10}, among
  others.} A researcher could impose stability on both models by using
a recursive window or relax stability for both by using a rolling
window; both approaches should not affect the test's size, but may
affect power.  But the researcher could instead impose stability on
the benchmark and relax it for the alternative by using a recursive
window for the benchmark and a rolling window for the alternative
model.  This approach could have a power advantage and is similar in
spirit to using a likelihood ratio test instead of an \textsc{lm} or
Wald test.

This statistic has a substantial advantage over existing \oos\
statistics for nested models: the alternative can be essentially
arbitrary, as long as it obeys the necessary moment conditions.  In
particular, researchers can use model selection techniques like the
\aic\ or \bic\ to determine the number of lags to include, the
particular exogenous variables to include, etc.  Other methods that
test a similar hypothesis are unable to handle these models
\citep[except][which does not allow the benchmark to be
estimated]{ClW:06}; \citet{GiW:06} are able to handle such models for
both the alternative and the benchmark, but they test a different
aspect of forecasting performance.

This paper focuses on nested models, as they have received the most
attention in the empirical and theoretical literature, but the
statistic can be used with non-nested models as well.  This generality
is useful, since \possessivecite{Wes:96} results do not apply to
non-nested models if they both encompass the true \dgp, which is
allowable under the null, and so the naive \oos\ $t$-test is not
asymptotically normal (in the limit, both models will converge to the
\dgp\ and give identical foreasts).  Constructing an \oos\ test for
non-nested models using \possessivecite{ClM:01} approach would be
difficult.  To my knowledge, this paper gives the first \oos\ test
of equal predictive ability for non-nested models that does not
implicitly rule out both models encompassing the
\dgp.\footnote{Again, \citet{GiW:06} does not make this implicit
  assumption, but their statistic tests a different null hypothesis.}

The next section presents our theoretical result and
Section~\ref{sec:2} presents two simulations that compare our
statistic to \poscw\ original test.  Section~\ref{sec:3} concludes.

\section{Theoretical Result}\label{sec:1}
We'll present our results as a single theorem.  The conditions are
essentially the same in existing papers
\citep[e.g.][]{ClW:07,ClW:06,Wes:96,WeM:98,Mcc:00,GiW:06}, but with a
modified window scheme.  Our result follows from \citet{Wes:96},
treating the alternative model's forecasts as additional data. Note
that asymptotic equivalence does not hold so we need to account for
parameter estimation error.

\begin{thm}\label{thm:1}
  Suppose that we have two models $\hat{y}_{1t}$ and $\hat{y}_{2t}$ to
  forecast the variable $y_t$, and have observations for
  $t=1,\dots,T$.  Assume the following hold:
  \begin{enumerate}
  \item The benchmark forecast $\hat{y}_{1,t}$, is estimated using
    \ols\ with a recursive window: $\hat{y}_{1,t+1} =
    x_t'\hat{\beta}_t$ for some vector of predictors $x_t$ with
    \[\hat{\beta}_t = \Big(\sum_{s=1}^{t} x_{t-1} x_{t-1}'\Big)^{-1}
    \sum_{s=1}^t x_{t-1} y_t, \qquad t = R,\dots,T.\]
    Also define $\beta^{*} = \big(\sum_{t=1}^{T} \E x_{t-1} x_{t-1}'\big)^{-1}
    \sum_{t=1}^T \E x_{t-1} y_t$.
  \item The alternative forecast, $\hat{y}_{2,t}$, is estimated using
    a rolling window of fixed length $R$ ($R < T$), so $\hat{y}_{2,t+1} =
    \psi(y_t,z_t,\dots,y_{t-R+1}, z_{t-R+1})$ where $\psi$ is a known
    function and $z_t$ is a sequence of predictors that includes $x_t$.
  \item The series $y_t$ and $z_t$ are strong mixing of size
    $-r/(r-2)$ for $r>2$ and $y_t$, $\hat{y}_{2t}$, and $x_t$ have
    uniformly bounded $2 r+\delta$ moments for some positive $\delta$.
  \item Define \[f_t(\beta) = (y_{t+1} - x_t'\beta)^2 - (y_{t+1} -
    \hat{y}_{2,t+1})^2 + (x_t'\beta - \hat{y}_{2,t+1})^2\] and $\bar f
    = P^{-1} \sum_{t=R}^{T-1} f_t(\hat{\beta}_t)$. Then $f(\beta^{*})$
    is covariance stationary and $\bar f$ has uniformly positive long
    run variance.
  \end{enumerate}
  Under the null hypothesis that $y_t - \hat{y}_{1t}(\beta^{*})$ is a
  martingale difference sequence with respect to the filtration
  $\sigma(y_t, z_{t-1}, y_{t-1}, z_{t-2},\dots)$, $\sqrt{P} \bar f /
  \hat{\sigma} \to^d N(0,1)$, where $P = T-R$ and
  \begin{align*}
    \hat{\sigma}^2 &= \hat{S}_{ff} + 2 \hat{S}_{fh} (T^{-1}X'X)^{-1}
    \hat{F}' +
    2\hat{F} (T^{-1} X'X)^{-1} \hat{S}_{hh} (T^{-1}X'X)^{-1} \hat{F}'\\
    \hat{F} &= 2 P^{-1} \sum_{t=R}^{T-1} x_t (2 x_t'\hat{\beta}_t -
    y_{t+1} - \hat{y}_{2,t+1}) \\
    \hat{S}_{ff} &= P^{-1}\sum_{t=R}^{T-1} (f_t(\hat{\beta}_t) - \bar
    f)^2 \\
    \hat{S}_{fh} &= P^{-1} \sum_{t=R}^{T-1} (f_t(\hat{\beta}_t) -
    \bar{f})(h_t(\hat{\beta}_t) - \bar h)' \\
    \hat{S}_{hh} &= P^{-1} \sum_{t=R}^{T-1} (h_t(\hat{\beta}_t) - \bar h)(h_t(\hat{\beta}_t) - \bar h)' \\
    h_t(\beta) &= x_t(y_{t+1} - x_t'\beta) \\
    \intertext{and} \bar h &= P^{-1} \sum_{t=R}^{T-1}
    h_t(\hat{\beta}_t).
  \end{align*}
\end{thm}

\begin{proof}[Proof of Theorem \ref{thm:1}] Let $\tilde{R}_T$ satisfy
  $\tilde{R}_T \to \infty$ and $\tilde{R}_T/T \to 0$ and define
  $\tilde{P} = T - \tilde{R}$.  Then we can replace $P$ and $R$ with
  $\tilde{P}$ and $\tilde{R}$ in the preceding equations without
  changing the asymptotic distributions of the statistics. Note that
  $\tilde{P}/\tilde{R} \to \infty$. It is straightforward to verify
  that the assumptions of Theorem 4.1 of \citet{Wes:96} are satisfied
  (with the moment and mixing conditions weakened as in
  \citet{Mcc:00}).  Note that \[f_t(\beta^{*})= 2 (y_{t+1} -
  x_t'\beta^{*})(x_t'\beta^{*} - \hat{y}_{2,t+1}) \quad \text{a.s.}\]
  under the null, as in \citet{ClW:07}, so we do not need to use a
  \hac\ estimator of the variance.
\end{proof}
The following remarks are relevant to Theorem~\ref{thm:1}:

\begin{rem}
  Forecasters will almost always be interested in the one-sided
  alternative that $\E f_t(\beta^{*}) > 0$; i.e. that the alternative
  model is expected to forecast better than the benchmark.
\end{rem}

@@

\begin{rem}
  The requirement that the asymptotic variance of $\bar f$ be
  uniformly positive is much less restrictive than in \cite{Wes:96}.
  As in \cite{GiW:06} and \citet{ClW:06,ClW:07}, the assumption only
  serves to rule out pathological cases---for example, letting both
  the benchmark and the alternative model be white noise. In
  \citet{Wes:96}, this assumption is a restriction on the \dgp\ as
  well as the forecasting models, but in this paper it is a
  restriction only on the models.
\end{rem}

\begin{rem}
  The statistic we present tests the null hypothesis that the forecast
  errors from the population version of the benchmark model are a
  martingale difference sequence.  This may not be appropriate,
  depending on the loss function or utility function of interest.  Our
  statistic can be modified to test implications of optimal forecasts
  under other loss functions \citep[see][]{PaT:07,PaT:07b}.
\end{rem}

\section{Monte Carlo Results}\label{sec:2}
\input{mc}
\section{Conclusion}\label{sec:3}
In this paper, we present an \oos\ test statistic similar to \poscw\
that is asymptotically normal when comparing nested models.  We do so
by estimating the alternative model using a fixed-length rolling
window---as Clark and West do---but estimating the benchmark model
with a recurisve window.  We also present simulations that indicate
the new statistic behaves similarly to Clark and West's original test,
suggesting that the new statistic is a suitable replacement in applied
research.
